Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-05-12 paper.qmd
Updated discussion

Our algorithm can also be useful in practice to decide on a reasonable NLDR layout that accurately represents the data, as described with a single-cell gene expression data, **PBMC3k**. The published results of UMAP with default (hyper-)parameter setting by @chen2024 were evaluated. The data showed three clusters, two of which were close. But the published layout showed three well-separated clusters. Therefore, by changing the (hyper-)parameter, we found a better layout with UMAP, which is more representative the data. Additionally, the model discovers non-uniform data distribution and non-linear structures within the clusters that are not visible in the UMAP layout, demonstrating the ability of our model in uncovering hidden data characteristics. Also, as explained in Appendix, with recently introduced approach, scDEED also proved the layout suggested by our layout is better than the published one.

>>>>

Our algorithm can also be useful in practice to decide on a reasonable NLDR layout that accurately represents the data, as described with a single-cell gene expression data, **PBMC3k**. The published results of UMAP with default (hyper-)parameter setting by @chen2024 were evaluated. The data showed three close clusters. But the published layout showed three well-separated clusters. Therefore, by changing the NLDR method to tSNE with default (hyper-)parameters, we found a better layout, which is more representative the data. Additionally, the model discovers non-uniform data distribution and non-linear structures within the clusters that are not visible in the \gD{} layout, demonstrating the ability of our model in uncovering hidden data characteristics. Also, as explained in Appendix, with recently introduced approach, scDEED also proved the layout suggested by our layout is better than the published one.
